Do investors in clean energy ETFs herd? The role of climate risks

### 2025-09-25 alt_energy_herding.qmd
addressed reviewer comments

Climate sustainability is a major concern for financial markets and investors. The growing interest in climate sustainability is primarily driven by the increasing materialisation of climate risks, and the actions of governments, institutions and organizations towards a sustainable future [@giglio2021a]. To safeguard returns, investors increasingly seek to hedge against climate risks by investing in green financial products. Although the evidence from the existing literature is mixed, returns from green financial products are somewhat comparable to traditional financial products [see amongst others, @decclesia2024; @nguyen2025; @pastor2022; and @naqvi2022]. In this new climate sustainability paradigm, investors face many pressures that not only bear on returns but also the stability of financial markets. For example, resulting regulations aimed at reducing emissions can surprisingly reduce the profitability of fossil-fuel-based companies, or the possible mispricing of assets from ignoring climate risks can lead to significant losses [@nguyen2025]. In addition to climate risks, a general change in investor attitudes can drive the inclusion of green assets in their portfolio can lead to systemic risk.^[These pressures notwithstanding the possible contribution that financial markets can play in mitigating and reducing the negative effects of climate change [@giglio2021a]] Specifically, climate change presents risks to the global financial system, and ultimately to investment portfolios, through two primary sources - physical and transition risks. Physical risks or direct impact refer to extreme climate events such as floods and droughts, which impact business operations and infrastructure; and transition risks are the policy, technological, and other costs that societies bear to achieve low carbon economies [@nguyen2025; and @giglio2021a amongst others]. Investors, therefore, recognise these risks and seek to mitigate them, as they seek return-enhancing green financial products.

@@

@@

where $R_{i,t}$ is the return and $R_{m,t}$ is the cross sectional average of returns for the sample of ETFs available for each day. The return dispersion measures the directional similarity of ETF returns to the market return. This return similarity forms the basis for the herding behaviour tests. Following @galariotis2015 we estimate @eq-herding:

@@

where $D^{herd}$ takes a value of 1 during periods of statistically significant herding (i.e., for days when the rolling t-statistic on $\gamma_2 < −1.96$ in @fig-rolling_window) and zero otherwise. $TRI$ is the transitional risk index and $PRI$ is the physical risk index. For details on the construction of $TRI$ and $PRI$, the reader can refer to the paper of @bua2024.

>>>>

Climate sustainability is a major concern for financial markets and investors. The growing interest in climate sustainability is primarily driven by the increasing materialisation of climate risks, and the actions of governments, institutions and organizations towards a sustainable future [@giglio2021a]. To safeguard returns, investors increasingly seek to hedge against climate risks by investing in green financial products [@nguyen2025]. Although the evidence from the existing literature is mixed, returns from green financial products are somewhat comparable to traditional financial products [see amongst others, @decclesia2024; @pastor2022; and @naqvi2022]. In this new climate sustainability paradigm, investors face many pressures that not only bear on returns but also the stability of financial markets. For example, resulting regulations aimed at reducing emissions can surprisingly reduce the profitability of fossil-fuel-based companies, or the possible mispricing of assets from ignoring climate risks can lead to significant losses [@nguyen2025]. In addition to climate risks, a general change in investor attitudes can drive the inclusion of green assets in their portfolio can lead to systemic risk.^[These pressures notwithstanding the possible contribution that financial markets can play in mitigating and reducing the negative effects of climate change [@giglio2021a]] Specifically, climate change presents risks to the global financial system, and ultimately to investment portfolios, through two primary sources - physical and transition risks. Physical risks or direct impact refer to extreme climate events such as floods and droughts, which impact business operations and infrastructure; and transition risks are the policy, technological, and other costs that societies bear to achieve low carbon economies [@nguyen2025; and @giglio2021a amongst others]. Investors, therefore, recognise these risks and seek to mitigate them, as they seek return-enhancing green financial products.

@@

 We adopt the popular definition herding behaviour, which follows from the broader behaviour economics literature. In recent years behavioural economics showed that human nature is prone to, amongst others, emotional bias and heuristics [@aggarwal2014animal; @kahneman2003maps]. These biases can lead to inefficient decision making, which can at times result in market failures [@devenow1996]. In this context, herding behaviour is imitative behivour or correlated trades by financial market agents that can result in market volatility and crises when the market is not well informed. This is why we are interested in herding behaviour in the clean energy market.

@@

where $R_{i,t}$ is the return and $R_{m,t}$ is the cross sectional average of returns for the sample of ETFs available for each day. The return dispersion measures the directional similarity of ETF returns to the market return. This return similarity forms the basis for the herding behaviour tests. Next we estimate @eq-herding as follows:

@@

where $D^{herd}$ takes a value of 1 during periods of statistically significant herding (i.e., for days when the rolling t-statistic on $\gamma_2 < −1.96$ in @fig-rolling_window) and zero otherwise. $TRI$ is the transitional risk index and $PRI$ is the physical risk index^[These indexes were constructed using the text-based approach of @engle2020hedging. This approach compares authorative text on climate change with European news sourced from Reuters - assuming that climate change events are covered in the news. The final sample had 1 million news articles from January 2005 to October 2021. @Bua2024 aggregate the articles to a daily frequency, thereby creating daily measures of $TRI$ and $PRI$]. For full details on the construction of $TRI$ and $PRI$, the reader can refer to the paper of @bua2024.

### 2025-09-25 alt_energy_herding.tex
addressed reviewer comments

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[round]{natbib}
\bibliographystyle{apalike}

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
% \usepackage{hyperref}
% \hypersetup{
%     colorlinks,
%     linkcolor={blue!100!black},
%     citecolor={blue!100!black},
%     urlcolor={blue!100!black}
% }
% \usepackage{apacite}
% \usepackage[round]{natbib} 
% \usepackage{graphicx}
% \usepackage{float}
% \usepackage{caption}
% \usepackage[toc,page]{appendix}
% \usepackage{booktabs,caption}
% \usepackage[flushleft]{threeparttable}
% \usepackage{tabularx}
\usepackage{sfmath}
\usepackage{fontspec}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{amsfonts}
\usepackage{amsmath}
% \usepackage{xcolor}
% \usepackage{scrextend}
% \deffootnote[2em]{2em}{1em}{\textsuperscript{\thefootnotemark}\,}
% \newcolumntype{Y}{>{\centering\arraybackslash}X}
% \usepackage[a4paper]{geometry}
% \usepackage{caption}
% \usepackage[bottom,flushmargin,hang,multiple]{footmisc}
% \usepackage{pdflscape}
% \usepackage[paper=portrait,pagesize]{typearea}
% \usepackage{rotating}
% \usepackage{fullpage}
% \usepackage{tabu}
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{rotating}
\newcommand{\bsideways}{\begin{sidewaystable}[htbp]}
\newcommand{\esideways}{\end{sidewaystable}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={blue!100!black},
  filecolor={Maroon},
  citecolor={blue!100!black},
  urlcolor={blue!100!black},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}
\begin{document}

@@

Climate sustainability is a major concern for financial markets and
investors. The growing interest in climate sustainability is primarily
driven by the increasing materialisation of climate risks, and the
actions of governments, institutions and organizations towards a
sustainable future \citep{giglio2021a}. To safeguard returns, investors
increasingly seek to hedge against climate risks by investing in green
financial products. Although the evidence from the existing literature
is mixed, returns from green financial products are somewhat comparable
to traditional financial products \citetext{\citealp[see amongst
others,][]{decclesia2024}; \citealp{nguyen2025}; \citealp{pastor2022}; \citealp[and][]{naqvi2022}}.
In this new climate sustainability paradigm, investors face many
pressures that not only bear on returns but also the stability of
financial markets. For example, resulting regulations aimed at reducing
emissions can surprisingly reduce the profitability of fossil-fuel-based
companies, or the possible mispricing of assets from ignoring climate
risks can lead to significant losses \citep{nguyen2025}. In addition to
climate risks, a general change in investor attitudes can drive the
inclusion of green assets in their portfolio can lead to systemic
risk.\footnote{These pressures notwithstanding the possible contribution
  that financial markets can play in mitigating and reducing the
  negative effects of climate change \citep{giglio2021a}} Specifically,
climate change presents risks to the global financial system, and
ultimately to investment portfolios, through two primary sources -
physical and transition risks. Physical risks or direct impact refer to
extreme climate events such as floods and droughts, which impact
business operations and infrastructure; and transition risks are the
policy, technological, and other costs that societies bear to achieve
low carbon economies \citetext{\citealp{nguyen2025}; \citealp[and][
amongst others]{giglio2021a}}. Investors, therefore, recognise these
risks and seek to mitigate them, as they seek return-enhancing green
financial products.

@@

This study investigates herding behaviour in alternative (clean) energy
ETFs in the US between May 1 2016 and June 19 2024, showing evidence of
significant herding that is asymmetric and time-varying. We then study
whether climate-related uncertainty can affect the herding behaviour in
the US clean energy ETFs market. Methodologically, we follow the
standard herding tests by \citet{christie1995} and \citet{chang2000}.
Notably, we supplement the traditional approach with quantile
regressions \citep{koenker1978} in order to capture how herding differs
across various quantiles of the returns dispersion and in up and down
markets. Lastly, to establish a link between climate risk and herding
behaviour, we differentiating between transitional and physical risks
\citep{bua2024} and provide valuable insights on their potential impact
on the likelihood of herding in clean energy ETFs.

@@

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_files/figure-pdf/fig-csad-1.pdf}}

@@

where \(R_{i,t}\) is the return and \(R_{m,t}\) is the cross sectional
average of returns for the sample of ETFs available for each day. The
return dispersion measures the directional similarity of ETF returns to
the market return. This return similarity forms the basis for the
herding behaviour tests. Following \citet{galariotis2015} we estimate
Equation~\ref{eq-herding}:

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_files/figure-pdf/fig-rolling_window-1.pdf}}

@@

where \(D^{herd}\) takes a value of 1 during periods of statistically
significant herding (i.e., for days when the rolling t-statistic on
\(\gamma_2 < −1.96\) in Figure~\ref{fig-rolling_window}) and zero
otherwise. \(TRI\) is the transitional risk index and \(PRI\) is the
physical risk index. For details on the construction of \(TRI\) and
\(PRI\), the reader can refer to the paper of \citet{bua2024}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
% \usepackage{hyperref}
% \hypersetup{
%     colorlinks,
%     linkcolor={blue!100!black},
%     citecolor={blue!100!black},
%     urlcolor={blue!100!black}
% }
% \usepackage{apacite}
% \usepackage[round]{natbib} 
% \usepackage{graphicx}
% \usepackage{float}
% \usepackage{caption}
% \usepackage[toc,page]{appendix}
% \usepackage{booktabs,caption}
% \usepackage[flushleft]{threeparttable}
% \usepackage{tabularx}
\usepackage{sfmath}
\usepackage{fontspec}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{amsfonts}
\usepackage{amsmath}
% \usepackage{xcolor}
% \usepackage{scrextend}
% \deffootnote[2em]{2em}{1em}{\textsuperscript{\thefootnotemark}\,}
% \newcolumntype{Y}{>{\centering\arraybackslash}X}
% \usepackage[a4paper]{geometry}
% \usepackage{caption}
% \usepackage[bottom,flushmargin,hang,multiple]{footmisc}
% \usepackage{pdflscape}
% \usepackage[paper=portrait,pagesize]{typearea}
% \usepackage{rotating}
% \usepackage{fullpage}
% \usepackage{tabu}
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{rotating}
\newcommand{\bsideways}{\begin{sidewaystable}[htbp]}
\newcommand{\esideways}{\end{sidewaystable}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[round]{natbib}
\bibliographystyle{apalike}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue!100!black},
  filecolor={Maroon},
  citecolor={blue!100!black},
  urlcolor={blue!100!black},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

@@

Climate sustainability is a major concern for financial markets and
investors. The growing interest in climate sustainability is primarily
driven by the increasing materialisation of climate risks, and the
actions of governments, institutions and organizations towards a
sustainable future \citep{giglio2021a}. To safeguard returns, investors
increasingly seek to hedge against climate risks by investing in green
financial products \citep{nguyen2025}. Although the evidence from the
existing literature is mixed, returns from green financial products are
somewhat comparable to traditional financial products
\citetext{\citealp[see amongst
others,][]{decclesia2024}; \citealp{pastor2022}; \citealp[and][]{naqvi2022}}.
In this new climate sustainability paradigm, investors face many
pressures that not only bear on returns but also the stability of
financial markets. For example, resulting regulations aimed at reducing
emissions can surprisingly reduce the profitability of fossil-fuel-based
companies, or the possible mispricing of assets from ignoring climate
risks can lead to significant losses \citep{nguyen2025}. In addition to
climate risks, a general change in investor attitudes can drive the
inclusion of green assets in their portfolio can lead to systemic
risk.\footnote{These pressures notwithstanding the possible contribution
  that financial markets can play in mitigating and reducing the
  negative effects of climate change \citep{giglio2021a}} Specifically,
climate change presents risks to the global financial system, and
ultimately to investment portfolios, through two primary sources -
physical and transition risks. Physical risks or direct impact refer to
extreme climate events such as floods and droughts, which impact
business operations and infrastructure; and transition risks are the
policy, technological, and other costs that societies bear to achieve
low carbon economies \citetext{\citealp{nguyen2025}; \citealp[and][
amongst others]{giglio2021a}}. Investors, therefore, recognise these
risks and seek to mitigate them, as they seek return-enhancing green
financial products.

@@

This study investigates herding behaviour in alternative (clean) energy
ETFs in the US between May 1 2016 and June 19 2024, showing evidence of
significant herding that is asymmetric and time-varying. We then study
whether climate-related uncertainty can affect the herding behaviour in
the US clean energy ETFs market. Methodologically, we follow the
standard herding tests by \citet{christie1995} and \citet{chang2000}.
Notably, we supplement the traditional approach with quantile
regressions \citep[ ]{koenker1978} in order to capture how herding
differs across various quantiles of the returns dispersion and in up and
down markets. Lastly, to establish a link between climate risk and
herding behaviour, we differentiating between transitional and physical
risks \citep{bua2024} and provide valuable insights on their potential
impact on the likelihood of herding in clean energy ETFs.

@@

We adopt the popular definition herding behaviour, which follows from
the broader behaviour economics literature. In recent years behavioural
economics showed that human nature is prone to, amongst others,
emotional bias and heuristics
\citep{aggarwal2014animal, kahneman2003maps}. These biases can lead to
inefficient decision making, which can at times result in market
failures \citep{devenow1996}. In this context, herding behaviour is
imitative behivour or correlated trades by financial market agents that
can result in market volatility and crises when the market is not well
informed. This is why we are interested in herding behaviour in the
clean energy market.

@@

\includegraphics{alt_energy_herding_files/figure-pdf/fig-csad-1.pdf}

@@

where \(R_{i,t}\) is the return and \(R_{m,t}\) is the cross sectional
average of returns for the sample of ETFs available for each day. The
return dispersion measures the directional similarity of ETF returns to
the market return. This return similarity forms the basis for the
herding behaviour tests. Next we estimate Equation~\ref{eq-herding} as
follows:

@@

\includegraphics{alt_energy_herding_files/figure-pdf/fig-rolling_window-1.pdf}

@@

where \(D^{herd}\) takes a value of 1 during periods of statistically
significant herding (i.e., for days when the rolling t-statistic on
\(\gamma_2 < −1.96\) in Figure~\ref{fig-rolling_window}) and zero
otherwise. \(TRI\) is the transitional risk index and \(PRI\) is the
physical risk index\footnote{These indexes were constructed using the
  text-based approach of \citet{engle2020hedging}. This approach
  compares authorative text on climate change with European news sourced
  from Reuters - assuming that climate change events are covered in the
  news. The final sample had 1 million news articles from January 2005
  to October 2021. \citet{Bua2024} aggregate the articles to a daily
  frequency, thereby creating daily measures of \(TRI\) and \(PRI\)}.
For full details on the construction of \(TRI\) and \(PRI\), the reader
can refer to the paper of \citet{bua2024}.
